---
title: "Eval Landsharc sims"
author: "Katie Lotterhos"
date: "May 24, 2016"
output: html_document
---
### Load packages and data
```{r setup, include=FALSE}
#setwd("/Users/katie/Desktop/MINOTAUReval")
source("evalsims/ComparePlot2.R")
source("evalsims/distanceFunctionsOther.R")
source("evalsims/getEmpiricalP3_BH.R")
source("evalsims/Getdf.R")


if(!("devtools" %in% installed.packages())){install.packages("devtools", dependencies=TRUE)}
if(!("rrcovNA" %in% installed.packages())){install.packages("rrcovNA")}

library(qvalue)
library(rrcovNA)
library(devtools)
install_github("NESCent/MINOTAUR")#, build_vignettes=TRUE)
library(MINOTAUR)

headdir <- "data/KatieSims/"
resultsdir <- "~/Desktop/MINOTAUReval/results/"
filelist <- list.files(headdir)
filelist <- filelist[-55]
length(filelist)
if (!length(filelist)==72){print("Errr: missing simulation files")}

sessionInfo()
```

### Loop through each dataset
The following chunk of code does for each dataset:
1) calculate multivariate distances based on default covariance matrix
2) compares covariance matrix estimated from just neutral loci to covariance matrix on all data and MCD estimate of covariance
3) calculate multivariate distances based on MCD covariance matrix
4) calculates empirical power for all stats

```{r}


cov_summary <- data.frame(filename=NA, 
                Comparison1=NA, Comparison2=NA, 
                ScatterDiff=NA, LocationDiff=NA)
head(cov_summary)

for (i in 1:length(filelist)){
  print(c(i, filelist[i]))
  
  infile <- paste(headdir, filelist[i], sep="")
  dat <- read.table(infile, header=TRUE)
  dat2 <- dat[dat$UseSNP,]
  cols <- which(names(dat) %in% c("rho", "log.bf", "xtx", "TW.Zscore"))

  # Covariance of all loci
    (Sall <- cov(dat2[,cols]))
    (Sall_mean <- colMeans(dat2[,cols]))
    
  # Covariance of neutral loci
    (Sneut <- cov(dat2[dat2$s_high==0,cols]))
    (Sneut_mean <- colMeans(dat2[dat2$s_high==0,cols]))
        
  # Covariance of MCD 75
    S_mcd75 <- CovNAMcd(dat2[,cols], alpha=0.75)
    S_mcd75@cov
    S_mcd75@center
    #S_mcd75
  # Covariance of MCD 99
    S_mcd99 <- CovNAMcd(dat2[,cols], alpha=0.99)
    S_mcd99@cov
    S_mcd99@center
    #S_mcd99    
    
    ### Comparisons with neutral
    cov_summary<- rbind(cov_summary, data.frame(
      filename=filelist[i],
      Comparison1= "Neutral",
      Comparison2= "All",
      ScatterDiff = mean((Sall-Sneut)^2),
      LocationDiff = mean((Sall_mean-Sneut_mean)^2)
    ))
    cov_summary<-rbind(cov_summary, data.frame(
      filename=filelist[i],
      Comparison1= "Neutral",
      Comparison2= "MCD 0.75",
      ScatterDiff= mean((Sneut-S_mcd75@cov)^2),
      LocationDiff= mean((Sneut_mean-S_mcd75@center)^2)
    ))
    cov_summary<-rbind(cov_summary, data.frame(
      filename=filelist[i],
      Comparison1= "Neutral",
      Comparison2= "MCD 0.99",
      ScatterDiff= mean((Sneut-S_mcd99@cov)^2),
      LocationDiff= mean((Sneut_mean-S_mcd99@center)^2)
    ))
    
     ### Comparisons with All
    cov_summary<- rbind(cov_summary, data.frame(
      filename=filelist[i],
      Comparison1= "All",
      Comparison2= "Neutral",
      ScatterDiff = mean((Sall-Sneut)^2),
      LocationDiff = mean((Sall_mean-Sneut_mean)^2)
    )) #this is redundant, for plotting only
    cov_summary<-rbind(cov_summary, data.frame(
      filename=filelist[i],
      Comparison1= "All",
      Comparison2= "MCD 0.75",
      ScatterDiff= mean((Sall-S_mcd75@cov)^2),
      LocationDiff= mean((Sall_mean-S_mcd75@center)^2)
    ))
    cov_summary<-rbind(cov_summary, data.frame(
      filename=filelist[i],
      Comparison1= "All",
      Comparison2= "MCD 0.99",
      ScatterDiff= mean((Sall-S_mcd99@cov)^2),
      LocationDiff= mean((Sall_mean-S_mcd99@center)^2)
    ))
}    
   write.csv(cov_summary, file = paste("results/covsummaryLandsharc.csv", sep=""))
``` 

```{r}
summarydf <- NULL
alpha=c(0.75, 0.99)
  # This is the proportion of points used in the MCD calculation

for (j in seq_along(alpha)){
  for (i in 1:length(filelist)){
    print(c(j, i, filelist[i]))
    infile <- paste(headdir, filelist[i], sep="")
    outfile <- paste(resultsdir, filelist[i], sep="")
    dat <- read.table(infile, header=TRUE)
    dat2 <- dat[dat$UseSNP,]
    
    ### Get stats based on covmat of all the data
    dat.out.defaultS <- Getdf(dat2, cols)
    colnames(dat.out.defaultS)[(ncol(dat.out.defaultS)-5):
                                 ncol(dat.out.defaultS)] <-
      c("pcs.default","Hcd.default", "Md.default", 
        "Hd.default", "Kd.default", "Nd.default")
    
    #head(dat.out.defaultS)
    #dim(dat.out)
      
      # Covariance of all loci
      Sall <- cov(dat2[,cols])
      
      # Covariance of MCD
      S_mcd <- CovNAMcd(dat2[,cols], alpha=alpha[j])
      #sum(S_mcd2@raw.wt) # this should be alpha*number rows in dat2
      
      # Compare cov of neutral data to cov_mcd of all data
      neut <- as.numeric(log(abs(Sneut)))
      all <- as.numeric(log(abs(Sall)))
      mcd <- as.numeric(log(abs(S_mcd@cov)))
      
      # Calculate the proportion of neutral/selected loci identified as robust (non-outliers)
      ProportionNeutralMCDRobust <- sum(dat2$s_high[S_mcd@raw.wt==1]==0)/sum(S_mcd@raw.wt==1)
      ProportionS0.005MCDRobust <- sum(dat2$s_high[S_mcd@raw.wt==1]==0.005)/sum(S_mcd@raw.wt==1)
      ProportionS0.01MCDRobust <- sum(dat2$s_high[S_mcd@raw.wt==1]==0.01)/sum(S_mcd@raw.wt==1)
      ProportionS0.1MCDRobust <- sum(dat2$s_high[S_mcd@raw.wt==1]==0.1)/sum(S_mcd@raw.wt==1)
      
      
      #par(mar=c(4,3,1,1), bty="n")
      # plot(neut, all, pch=19, col="blue",
      #      xlab="log(abs(covariance)):\nNeutral loci only", ylab="") 
      #  abline(0,1, lwd=2)
      #  abline(lm(all~neut), col="blue")
      # points(neut, mcd, pch=17, col="orange")
      # abline(lm(mcd~neut), col="orange")
      # points(neut, mcd, col="red")
      # abline(lm(mcd~neut), col="red", lwd=2)
      # legend(0,-2, c("raw", "mcd alpha=50", "mcd alpha=75"), col=c("blue", "orange", "red"), pch=c(19, 17, 1), bty="n")
       
      ### Get stats based on mcd covmat
      dat.out <- Getdf(dat.out.defaultS,cols,S=S_mcd@cov, M=S_mcd@center)
      head(dat.out)
      colnames(dat.out)[(ncol(dat.out)-3):
                                 ncol(dat.out)] <-
      c( "Md.mcd", "Hd.mcd", "Kd.mcd", "Nd.mcd") 
       head(dat.out)
       
      ### Get stats based on mcd covmat and robust points
      dat.out.subset <- Getdf(dat.out, cols, S=S_mcd@cov,  M=S_mcd@center, subset=which(S_mcd@raw.wt==1))
      head(dat.out.subset)
      colnames(dat.out.subset)[(ncol(dat.out.subset)-3):
                                 ncol(dat.out.subset)] <-
      c( "Md.mcd.subset", "Hd.mcd.subset", "Kd.mcd.subset", "Nd.mcd.subset") 
       head(dat.out.subset)
  
      dat.out <- dat.out.subset
    ### Make plots ####
  #    pdf(file = paste(outfile, ".pdf", sep=""),width = 4, height=6)
  #      par(mfrow=c(2,1), mar=c(4,4,1,1), oma=c(0,0,2,1))
  #      plot(dat.out[,cols[1]], dat.out[,cols[2]], col=factor(dat.out$s_high), xlab="Spearman's rho (GEA)", ylab="XTX (Fst analog)")
  #      plot(dat.out[,cols[1]], dat.out[,cols[3]], col=factor(dat.out$s_high), xlab="Spearman's rho (GEA)", ylab="Z-score (LFMM, GEA)")
  #      mtext(dat.out$demog[1] , side = 3, outer=TRUE)
  #    dev.off()
  #    pdf(file = paste(outfile, "mulitD.pdf", sep=""),width = 6, height=8)
  #      ComparePlot(dat.out, colorVect=factor(dat$s_high), 9500:9996)
  #    dev.off()
  
    ### Calculate Empirical Power ###
      out <- data.frame(infile = as.character(infile),
               demog = as.character(dat.out$demog[1]),
               xtx.ep = getEmpPower(dat.out$xtx, dat.out$s_high==0),
               bf.ep = getEmpPower(dat.out$log.bf, dat.out$s_high==0),
               rho.ep = getEmpPower(dat.out$rho, dat.out$s_high==0),
               lfmm.ep = getEmpPower(dat.out$TW.Zscore, dat.out$s_high==0),
               Hcd.ep = getEmpPower(dat.out$Hcd, dat.out$s_high==0),
               pcs.ep = getEmpPower(dat.out$pcs, dat.out$s_high==0),
               alpha = alpha[j],
               
               Md.default.ep = getEmpPower(dat.out$Md.default, dat.out$s_high==0),
               Hd.default.ep = getEmpPower(dat.out$Hd.default, dat.out$s_high==0),
               Kd.default.ep = getEmpPower(dat.out$Kd.default, dat.out$s_high==0),
               Nd.default.ep = getEmpPower(dat.out$Nd.default, dat.out$s_high==0),
               
               Md.mcd.ep = getEmpPower(dat.out$Md.mcd, dat.out$s_high==0),
               Hd.mcd.ep = getEmpPower(dat.out$Hd.mcd, dat.out$s_high==0),
               Kd.mcd.ep = getEmpPower(dat.out$Kd.mcd, dat.out$s_high==0),
               Nd.mcd.ep = getEmpPower(dat.out$Nd.mcd, dat.out$s_high==0),
                
               Md.mcd.subset.ep = getEmpPower(dat.out$Md.mcd.subset, dat.out$s_high==0),
               Hd.mcd.subset.ep = getEmpPower(dat.out$Hd.mcd.subset, dat.out$s_high==0),
               Kd.mcd.subset.ep = getEmpPower(dat.out$Kd.mcd.subset, dat.out$s_high==0),
               Nd.mcd.subset.ep = getEmpPower(dat.out$Nd.mcd.subset, dat.out$s_high==0),
               
              Md.mcd.subset.fp = sum(p.adjust(1-returnEmpP(dat.out$Md.mcd.subset[dat.out$s_high==0], dat.out$Md.mcd.subset[S_mcd@raw.wt==1 & dat.out$s_high==0]),method="BH")<0.05)/9900,
               
               Neut.cov.vs.All = mean(abs(neut-all)),
               Neut.cov.vs.MCD = mean(abs(neut-mcd)),
               
               ProportionNeutralMCDRobust,
               ProportionS0.005MCDRobust,
               ProportionS0.01MCDRobust,
               ProportionS0.1MCDRobust
      )
    if (i==1 & j==1){
      write.table(out, file = paste(resultsdir,"LandsharcSummary2.txt", sep=""), append=FALSE, col.names=TRUE, row.names=FALSE)
    }else{
        write.table(out, file = paste(resultsdir,"LandsharcSummary2.txt",sep=""), append=TRUE, col.names=FALSE, row.names=FALSE) 
                  }
    #summarydf <- rbind(summarydf,out)
  } # end loop through files
} # end loop through alpha
#write.table(summarydf,file = , sep=""))
```
