---
title: "Test Landsharc sims - make plots"
author: "Katie Lotterhos"
date: "May 24, 2016"
output: html_document
---

```{r setup, include=FALSE}

### Make summary plots
```{r}

summarydf <- read.table("results/LandsharcSummary.txt", header=TRUE)
colnames(summarydf)<-sub(".ep", "",colnames(summarydf))
head(summarydf)

pdf("results/LandsharcSummary.pdf", width = 15, height = 12)
    par(mfrow=c(4,1), mar=c(1,3,1,1),oma=c(1,3,1,0), cex=1.0)
    x <- 0.5
    y <- 1.1
    stats <- c(3:6,9:12,13:16)
    colors <- c(rep("magenta",4), rep("lightblue",4), rep("blue",4))
    boxplot(summarydf[summarydf$demog=="IM",stats], col=colors,
            ylim=c(0,1.2))
    text(x,y, "IM", cex=2)
    boxplot(summarydf[summarydf$demog=="IBD",stats], col=colors,
            ylim=c(0,1.2))
    text(x,y, "IBD", cex=2)
    boxplot(summarydf[summarydf$demog=="1R",stats], col=colors,
            ylim=c(0,1.2))
    text(x,y, "1R", cex=2)
    boxplot(summarydf[summarydf$demog=="2R",stats], col=colors,
            ylim=c(0,1.2))
    text(x,y, "2R", cex=2)
    mtext("Empirical Power", side=2, outer=TRUE)
dev.off()
# colMeans(summarydf[summarydf$demog=="IM",3:10])
# colMeans(summarydf[summarydf$demog=="IBD",3:10])
# colMeans(summarydf[summarydf$demog=="1R",3:10])
# colMeans(summarydf[summarydf$demog=="2R",3:10])


pdf("results/LandsharcSummary_MdHd.pdf", width = 10, height =8)
    par(mfrow=c(4,1), mar=c(1,3,1,1),oma=c(1,3,1,0), cex=1.0)
    x <- 0.5
    y <- 1.1
    stats <- c(3:6,9:10,13:14)
    colors <- c(rep("magenta",4), rep("lightblue",2), rep("blue",2))
    boxplot(summarydf[summarydf$demog=="IM",stats], col=colors,
            ylim=c(0,1.2))
    text(x,y, "IM", cex=2)
    boxplot(summarydf[summarydf$demog=="IBD",stats], col=colors,
            ylim=c(0,1.2))
    text(x,y, "IBD", cex=2)
    boxplot(summarydf[summarydf$demog=="1R",stats], col=colors,
            ylim=c(0,1.2))
    text(x,y, "1R", cex=2)
    boxplot(summarydf[summarydf$demog=="2R",stats], col=colors,
            ylim=c(0,1.2))
    text(x,y, "2R", cex=2)
    mtext("Empirical Power", side=2, outer=TRUE)
dev.off()
```
