---
title: "Eval Landsharc sims"
author: "Katie Lotterhos"
date: "May 24, 2016"
output: html_document
---
### Load packages and data
```{r setup, include=FALSE}
#setwd("/Users/katie/Desktop/MINOTAUReval")
source("evalsims/ComparePlot2.R")
source("evalsims/distanceFunctionsOther.R")
source("evalsims/getEmpiricalP2.R")
source("evalsims/Getdf.R")


if(!("devtools" %in% installed.packages())){install.packages("devtools", dependencies=TRUE)}
if(!("rrcovNA" %in% installed.packages())){install.packages("rrcovNA")}
if(!("qvalue" %in% installed.packages())){
  source("https://bioconductor.org/biocLite.R")
  biocLite("qvalue")
  }

library(qvalue)
library(rrcovNA)
library(devtools)
install_github("NESCent/MINOTAUR")#, build_vignettes=TRUE)
library(MINOTAUR)

headdir <- "data/KatieSims/"
resultsdir <- "~/Desktop/MINOTAUReval/results/"
filelist <- list.files(headdir)
filelist <- filelist[-55]
length(filelist)
if (!length(filelist)==72){print("Errr: missing simulation files")}

sessionInfo()
```

### Loop through each dataset
The following chunk of code does for each dataset:
1) calculate multivariate distances based on default covariance matrix
2) compares covariance matrix estimated from just neutral loci to covariance matrix on all data and MCD estimate of covariance
3) calculate multivariate distances based on MCD covariance matrix
4) calculates empirical power for all stats

```{r}
summarydf <- NULL
for (i in 1:length(filelist)){
  print(c(i, filelist[i]))
  infile <- paste(headdir, filelist[i], sep="")
  outfile <- paste(resultsdir, filelist[i], sep="")
  dat <- read.table(infile, header=TRUE)
  cols <- which(names(dat) %in% c("rho", "log.bf", "xtx", "TW.Zscore"))
  dat2 <- dat[dat$UseSNP,]
  
  ### Get stats based on covmat of all the data
  dat.out.defaultS <- Getdf(dat2, cols)
  colnames(dat.out.defaultS)[(ncol(dat.out.defaultS)-5):
                               ncol(dat.out.defaultS)] <-
    c("pcs.default","Hcd.default", "Md.default", 
      "Hd.default", "Kd.default", "Nd.default")
  #head(dat.out.defaultS)
  #dim(dat.out)
  
    # Covariance of neutral loci
    Sneut <- cov(dat2[dat2$s_high==0,cols])
    
    # Covariance of all loci
    Sall <- cov(dat2[,cols])
    
    # Covariance of MCD
    S_mcd <- CovNAMcd(dat2[,cols])
    #length(S_mcd@best)
    #sum(S_mcd@raw.wt)
    S_mcd75 <- CovNAMcd(dat2[,cols], alpha=0.75)
    #length(S_mcd2@best)
    #sum(S_mcd2@raw.wt)
    
    # Compare cov of neutral data to cov_mcd of all data
    neut <- as.numeric(log(abs(Sneut)))
    all <- as.numeric(log(abs(Sall)))
    mcd <- as.numeric(log(abs(S_mcd@cov)))
    mcd.75 <- as.numeric(log(abs(S_mcd75@cov)))
    par(mar=c(4,3,1,1), bty="n")
     plot(neut, all, pch=19, col="blue",
          xlab="log(abs(covariance)):\nNeutral loci only", ylab="") 
      abline(0,1, lwd=2)
      abline(lm(all~neut), col="blue")
     points(neut, mcd, pch=17, col="orange")
     abline(lm(mcd~neut), col="orange")
     points(neut, mcd.75, col="red")
     abline(lm(mcd.75~neut), col="red", lwd=2)
     legend(0,-2, c("raw", "mcd alpha=50", "mcd alpha=75"), col=c("blue", "orange", "red"), pch=c(19, 17, 1), bty="n")
     
    ### Get stats based on mcd covmat
    dat.out <- Getdf(dat.out.defaultS,cols,S=S_mcd75@cov, M=S_mcd75@center)
    head(dat.out)
    colnames(dat.out)[(ncol(dat.out)-3):
                               ncol(dat.out)] <-
    c( "Md.mcd", "Hd.mcd", "Kd.mcd", "Nd.mcd") 
     head(dat.out)
     
    ### Get stats based on mcd covmat and robust points
    dat.out.subset <- Getdf(dat.out, cols, S=S_mcd75@cov,  M=S_mcd75@center, subset=which(S_mcd75@raw.wt==1))
    head(dat.out.subset)
    colnames(dat.out.subset)[(ncol(dat.out.subset)-3):
                               ncol(dat.out.subset)] <-
    c( "Md.mcd.subset", "Hd.mcd.subset", "Kd.mcd.subset", "Nd.mcd.subset") 
     head(dat.out.subset)

    dat.out <- dat.out.subset
  ### Make plots ####
#    pdf(file = paste(outfile, ".pdf", sep=""),width = 4, height=6)
#      par(mfrow=c(2,1), mar=c(4,4,1,1), oma=c(0,0,2,1))
#      plot(dat.out[,cols[1]], dat.out[,cols[2]], col=factor(dat.out$s_high), xlab="Spearman's rho (GEA)", ylab="XTX (Fst analog)")
#      plot(dat.out[,cols[1]], dat.out[,cols[3]], col=factor(dat.out$s_high), xlab="Spearman's rho (GEA)", ylab="Z-score (LFMM, GEA)")
#      mtext(dat.out$demog[1] , side = 3, outer=TRUE)
#    dev.off()
#    pdf(file = paste(outfile, "mulitD.pdf", sep=""),width = 6, height=8)
#      ComparePlot(dat.out, colorVect=factor(dat$s_high), 9500:9996)
#    dev.off()

  ### Calculate Empirical Power ###
    out <- data.frame(infile = as.character(infile),
             demog = as.character(dat.out$demog[1]),
             xtx.ep = getEmpPower(dat.out$xtx, dat.out$s_high==0),
             bf.ep = getEmpPower(dat.out$log.bf, dat.out$s_high==0),
             rho.ep = getEmpPower(dat.out$rho, dat.out$s_high==0),
             lfmm.ep = getEmpPower(dat.out$TW.Zscore, dat.out$s_high==0),
             Hcd.ep = getEmpPower(dat.out$Hcd, dat.out$s_high==0),
             pcs.ep = getEmpPower(dat.out$pcs, dat.out$s_high==0),
             
             Md.default.ep = getEmpPower(dat.out$Md.default, dat.out$s_high==0),
             Hd.default.ep = getEmpPower(dat.out$Hd.default, dat.out$s_high==0),
             Kd.default.ep = getEmpPower(dat.out$Kd.default, dat.out$s_high==0),
             Nd.default.ep = getEmpPower(dat.out$Nd.default, dat.out$s_high==0),
             
             Md.mcd.ep = getEmpPower(dat.out$Md.mcd, dat.out$s_high==0),
             Hd.mcd.ep = getEmpPower(dat.out$Hd.mcd, dat.out$s_high==0),
             Kd.mcd.ep = getEmpPower(dat.out$Kd.mcd, dat.out$s_high==0),
             Nd.mcd.ep = getEmpPower(dat.out$Nd.mcd, dat.out$s_high==0),
              
             Md.mcd.subset.ep = getEmpPower(dat.out$Md.mcd.subset, dat.out$s_high==0),
             Hd.mcd.subset.ep = getEmpPower(dat.out$Hd.mcd.subset, dat.out$s_high==0),
             Kd.mcd.subset.ep = getEmpPower(dat.out$Kd.mcd.subset, dat.out$s_high==0),
             Nd.mcd.subset.ep = getEmpPower(dat.out$Nd.mcd.subset, dat.out$s_high==0)           
    )
  summarydf <- rbind(summarydf,out)
}

write.table(summarydf,file = paste(resultsdir,"LandsharcSummary.txt", sep=""))
```
