---
title: "Test multivariate outliers and cov. calculations manuscript 1"
author: "Katie Lotterhos"
date: "April 11, 2016"
output: html_document
---

```{r}
#getwd() 
#
#setwd("/Users/katie/Desktop/MINOTAUReval/")
source("evalsims/distanceFunctionsOther.R")
#source("misc/evalsims/ComparePlot.R")
source("evalsims/Getdf.R")
source("evalsims/getEmpiricalP2.R")

if(!("devtools" %in% installed.packages())){install.packages("devtools", dependencies=TRUE)}
if(!("rrcovNA" %in% installed.packages())){install.packages("rrcovNA")}
if(!("qvalue" %in% installed.packages())){
  source("https://bioconductor.org/biocLite.R")
  biocLite("qvalue")
  }

library(qvalue)
library(rrcovNA)
library(devtools)
install_github("NESCent/MINOTAUR")#, build_vignettes=TRUE)
library(MINOTAUR)
```

#### Non Parameteric Simulation Example
```{r}
  np <- read.csv("data/df_inverse.csv")
  head(np)
  tail(np)
  nrow(np)
  np2 <- Getdf(np)
  col <- rep("grey", nrow(np2))
  pch <- rep(19, nrow(np2))
  cex <- rep(0.8, nrow(np2))
  col[1001:1002] <- "blue"
  pch[1001:1002] <- c(17,18)
  cex[1001:1002] <- 2
  ind <- c(1:1000/10000,.11,0.12)

png("figures_man1/nonPara_dist.png", width=4, height=4, res=300, units="in")
  plot(np, col=col, pch=pch, cex=cex)
dev.off()

  plot(np, col=col, pch=pch, cex=cex)


png("figures_man1/nonPara_log.png", width=10, height=12, res=300, units="in")
  make_nonPara_log <- function(){
  par(mfrow=c(4,2), mar=c(3,4,1,1), bty="l")
    plot(ind,log(np2$Md), col=col, pch=pch, ylab= "log(Mahalanobis)", cex=cex, xaxt="n")
      #abline(sort(log(dfv3.out$Md[dfv3.out$s==0]))[9900*0.999],0)
      #text(0,1.5, "A", cex=2)
    #plot(0,0, xaxt="n", yaxt="n")
    plot(density(np2$Md), main="")
  
      #abline(sort(log(dfv3.out$Hd[dfv3.out$s==0]))[9900*0.999],0)
      #text(0,1.5, "B", cex=2)
      #text(0,2.3,round(getEmpPower(dfv3.out$Hd,dfv3.out$s_high==0),2))
    plot(ind,log(np2$Hd), col=col, pch=pch, ylab= "log(Harmonic mean dist.)(cov)", cex=cex, xaxt="n")
     plot(density(np2$Hd), main="")


  
    plot(ind, log(np2$Kd), col=col, pch=pch, ylab= "log(Kernel density) (cov)", cex=cex, xaxt="n")
    plot(density(np2$Kd), main="") 

   
    plot(ind, log(np2$Nd), col=col, pch=pch, ylab= "log(Nearest neighbor)(cov)", cex=cex, xaxt="n")
    plot(density(np2$Nd), main="") 
}
dev.off()
make_nonPara_log()
```

### Two refugia example

Load data and output dataframe with two kinds of multivariate stats: one that only standardizes by the variance in each dimension, and one that standardizes by the variance-covariance matrix (for three functions: harmonicDist, neighborDist, and kernelDist.)  The `Getdf()` function uses kernelDeviance to calculate the best bandwidth.


```{r}
  d1 <- read.table("data/KatieSims/2R_R30_1351142970_988_6_NumPops=30_NumInd=20Bayenv2LFMMpca.Cpval", header=TRUE)
  head(d1)
  dfv <- d1[c(1,3,4,5,10,12,13,15:17,34)]
  dim(dfv)
  head(dfv)
  #dfv2 <- dfv[dfv$SNPIncluded,]
  colnums <- 8:11
  dfv <- dfv[complete.cases(dfv),]
  dim(dfv)
  #head(dfv)
  #str(dfv)
  #cbind(colnames(dfv))

  table(dfv$s_high)
  names(dfv)[colnums]

```

#### Get the covariance using all data (default)
```{r}
  dfv2.out <- Getdf(dfv, colnums)

  dim(dfv)
  dim(dfv2.out)
  head(dfv2.out)
  dfv3.out <- dfv2.out[,-c(12:13)]
  head(dfv3.out)
 
  head(dfv3.out)
```

#### Compare CovNAMcd to covariance using all data

Darren this is the code chunck you want

```{r}
  # Covariance matrix of all data
    S <- cov(dfv[!is.na(rowSums(dfv[,colnums])),colnums])
    S
    solve(S)
  # Covariance of neutral loci
    Sneut <- cov(dfv[dfv$s_high==0,colnums])
    Sneut
    solve(Sneut)
  par(mfrow=c(1,1), mar=c(4,4,1,1))
  plot(log(abs(Sneut)), log(abs(S))); abline(0,1)

  S_mcd <- CovNAMcd(dfv[,colnums])
  S_mcd
  Sneut_mcd <- CovNAMcd(dfv[dfv$s_high==0,colnums])
  Sneut_mcd 
  str(S_mcd@cov)
  
  length(S_mcd@best)
  sum(S_mcd@best>9900) # selected loci included
  sum(S_mcd@best<100) # example of neutral loci included
  
  # compare the MCD covariance to the covariance calculated on the "best" points
  plot(log(abs(S_mcd@cov)), log(abs(cov(dfv[S_mcd@best,colnums]))))#, xlim=c(0,10), ylim=c(0,10))
  abline(0,1)
    # not sure why they aren't the same
  
# Compare cov of neutral data to cov_mcd of all data
  plot(log(abs(Sneut)), log(abs(S_mcd@cov))); abline(0,1)

# Compare cov_mcd of all data to cov_mcd of neutral data
  plot(log(abs(Sneut_mcd@cov)), log(abs(S_mcd@cov))); abline(0,1)

  dfv4.out <- Getdf(dfv, colnums, S=S_mcd@cov,  M=S_mcd@center)
  head(dfv4.out)
  dfv5.out <- dfv4.out[,-c(12:13)]
#write.table(dfv5.out, "data/TwoRefSimForShinyM.txt",row.names=FALSE)

```

#### Compare to CovNAMcd and to subset points
```{r}
  dfv6.subset <- Getdf(dfv, colnums, S=S_mcd@cov, M=S_mcd@center, subset=S_mcd@best)
  dfv6.subset <- dfv6.subset[,-c(12:13)]
```


### Set color levels for plotting
```{r}
 col <- factor(dfv2.out$s_high)
  rgb(0,0,0,0.1)
  levels(col) = c("#0000001A",  "#9ad0f3", "#0072B2", "#D55E00")
  col <- as.character(col)
  ind <- c(1:9900/100, 101:199)
  cex <- c(rep(0.8, 9900), rep(1.1, 100))
  pch <- c(rep(19, 9900), rep(17, 100))
```

### Plot classical vs. robust

```{r}
#install.packages("cluster")
library(cluster)
library(MASS)

par(mfrow=c(1,1), mar=c(4,4,1,1))
plot(dfv[,colnums[2:3]], col=rgb(0,0,0,0.05), pch=pch, ylim=c(20,80), bty="n")
points(dfv[9901:9999,colnums[2:3]], col=col[9901:9999], pch=pch[9901:9999], ylim=c(20,80))
X <- dfv[,colnums[2:3]]

### Add classical covariance
  C.ls <- cov(X) ; m.ls <- colMeans(X)
  d2.99 <- qchisq(0.99, df = 2)
  lines(ellipsoidPoints(C.ls, d2.99, loc=m.ls), col="black", lwd=2)

### Add robust covariance to plot
  par(mfrow=c(1,1), mar=c(4,4,1,1))
plot(dfv[,colnums[2:3]], col=rgb(0,0,0,0.05), pch=pch, ylim=c(20,80), bty="n")
points(dfv[9901:9999,colnums[2:3]], col=col[9901:9999], pch=pch[9901:9999], ylim=c(20,80))
X <- dfv[,colnums[2:3]]
  C.ls <- cov(X) ; m.ls <- colMeans(X)
  d2.99 <- qchisq(0.99, df = 2)
  lines(ellipsoidPoints(C.ls, d2.99, loc=m.ls), col="black", lwd=2)
  Cxy <- cov.rob(X)
  str(Cxy)
  lines(ellipsoidPoints(Cxy$cov, d2 = d2.99, loc=Cxy$center), col="magenta", lwd=3)
  
### Add "best" points to plot
  par(mfrow=c(1,1), mar=c(4,4,1,1))
plot(dfv[,colnums[2:3]], col=rgb(0,0,0,0.05), pch=pch, ylim=c(20,80), bty="n")
points(dfv[9901:9999,colnums[2:3]], col=col[9901:9999], pch=pch[9901:9999], ylim=c(20,80))
X <- dfv[,colnums[2:3]]
  C.ls <- cov(X) ; m.ls <- colMeans(X)
  d2.99 <- qchisq(0.99, df = 2)
  lines(ellipsoidPoints(C.ls, d2.99, loc=m.ls), col="black", lwd=2)
  Cxy <- cov.rob(X)
  str(Cxy)
  lines(ellipsoidPoints(Cxy$cov, d2 = d2.99, loc=Cxy$center), col="magenta", lwd=3)
  points(X[Cxy$best,], pch=19, cex=0.1, col="magenta")  

  # add covariance based on "best" points
  par(mfrow=c(1,1), mar=c(4,4,1,1))
plot(dfv[,colnums[2:3]], col=rgb(0,0,0,0.05), pch=pch, ylim=c(20,80), bty="n")
points(dfv[9901:9999,colnums[2:3]], col=col[9901:9999], pch=pch[9901:9999], ylim=c(20,80))
X <- dfv[,colnums[2:3]]
  C.ls <- cov(X) ; m.ls <- colMeans(X)
  d2.99 <- qchisq(0.99, df = 2)
  lines(ellipsoidPoints(C.ls, d2.99, loc=m.ls), col="black", lwd=2)
  Cxy <- cov.rob(X)
  str(Cxy)
  lines(ellipsoidPoints(Cxy$cov, d2 = d2.99, loc=Cxy$center), col="magenta", lwd=3)
  points(X[Cxy$best,], pch=19, cex=0.1, col="magenta")  
  
  
    Cbest <- cov(X[Cxy$best,]); Mbest <- colMeans(X[Cxy$best,])
    lines(ellipsoidPoints(Cbest, d2 = d2.99, loc=Mbest), col="green", lwd=3)
    
      text(0.3,30, "All Data")
      text(0.1, 24, "Robust covariance", col = "Magenta")
      text(0.15, 39, "Cov of 'best' points", col="green")
```

### Compare cov.rob to CovNAMcd
```{r}
### Compare cov.rob function to mcd function
  S_rob <- cov.rob(dfv[,colnums])
  S_rob$cov
  cov(dfv[S_rob$best, colnums])
  identical(S_rob$best, S_mcd@best)
  S_rob2 <- cov.rob(dfv[,colnums])
  identical(S_rob$best, S_rob2$best)
    # every time is going to be a little different
  
  par(mfrow=c(3,1))
  plot(log(abs(S_mcd@cov)), log(abs(S_rob$cov)))
  abline(0,1)
  
  plot(log(abs(S_mcd@cov)), log(abs(cov(dfv[S_mcd@best,colnums]))))#, xlim=c(0,10), ylim=c(0,10))
  abline(0,1)
  
  plot(log(abs(S_rob$cov)), log(abs(cov(dfv[S_rob$best, colnums]))))
  abline(0,1)
```

### Plot univariate Distributions of Two Ref model

```{r}
  png("figures_man1/TwoRefUnivarDist.png", width=6, height=6, res=450, units="in")
  plot_TwoRefUnivarDist<-function(){
    dplot <- dfv[,colnums]
    dplot[,2] <- abs(dplot[,2]) #abs for rho
    names <- c("log(Bayes Factor)", "Spearman's rho", "XTX", "Z-Score")
    colnames(dplot)
    par(mfrow=c(3,3), mar=c(3,3,0,0), oma=c(1,1,1,1))
      # (1,1) BF vs rho
      plot(dplot[,1], dplot[,2], col=col, pch=pch, bty="n")
        mtext("Bayes Factor", side=1, line=2.5, cex=1)
        mtext("Spearman's rho", side=2, line=2.5, cex=1)
      # (1,2) XTX vs rho
      plot(dplot[,3], dplot[,2], col=col, pch=pch, bty="n")
      # (1,3) Z vs rhow
      plot(dplot[,4], dplot[,2], col=col, pch=pch, bty="n")
      # (2, 1) blank
      plot(NULL, NULL, xlim=c(0,1), ylim=c(0,1), xaxt="n", yaxt="n", bty="n")
      # (2,2) XTX vs BF
      plot(dplot[,3], dplot[,1], col=col, pch=pch, bty="n")
        mtext("XTX", side=1, line=2.5, cex=1)
        mtext("Bayes Factor", side=2, line=2.5, cex=1)
      # (2,3) Z vs BF
      plot(dplot[,4], dplot[,1], col=col, pch=pch, bty="n")
      # (3, 1) blank
      plot(NULL, NULL, xlim=c(0,1), ylim=c(0,1), xaxt="n", yaxt="n", bty="n")
      # (3, 2) blank
      plot(NULL, NULL, xlim=c(0,1), ylim=c(0,1), xaxt="n", yaxt="n", bty="n")
      # (2,3) Z vs XTX
      plot(dplot[,4], dplot[,3], col=col, pch=pch, bty="n")
          mtext("Z-score", side=1, line=2.5, cex=1)
          mtext("XTX", side=2, line=2.5, cex=1)
    }
    dev.off()
    
  plot_TwoRefUnivarDist()  
```

### Plot Two Ref example: compare default and mcd covariance ####

```{r, fig.height=8, fig.width=12}
#png("figures_man1/TwoRef_log_var.png", width=10, height=8, res=450, units="in")
makeplot <- function(dfv2.out){
   par(mfrow=c(2,2), mar=c(3,4,1,1), bty="l")
    plot(ind, log(dfv2.out$Md), col=col, pch=pch, ylab= "log(Mahalanobis)", cex=cex)
      abline(sort(log(dfv2.out$Md[dfv2.out$s==0]))[9900*0.999],0)
      text(5,max(log(dfv2.out$Md), na.rm=TRUE)*0.95, "A", cex=2)
      text(30,max(log(dfv2.out$Md), na.rm=TRUE)*0.95,round(getEmpPower(dfv2.out$Md,dfv2.out$s_high==0),2))
      legend(125, mean(log(dfv2.out$Md), na.rm=TRUE), legend=c("Neutral", "s = 0.005", "s=0.01", "s=0.1"),
             pch=c(19, 17, 17,17), col=c("grey",  "#9ad0f3", "#0072B2", "#D55E00"))

    plot(ind, log(dfv2.out$Hd), col=col, pch=pch, ylab= "log(Harmonic mean dist.)", cex=cex)
      abline(sort(log(dfv2.out$Hd[dfv2.out$s==0]))[9900*0.999],0)
      text(5,max(log(dfv2.out$Hd), na.rm=TRUE)*0.95, "B", cex=2)
      text(30,max(log(dfv2.out$Hd), na.rm=TRUE)*0.95,round(getEmpPower(dfv2.out$Hd,dfv2.out$s_high==0),2))

    plot(ind, log(dfv2.out$Kd), col=col, pch=pch, ylab= "log(Kernel density)", cex=cex)
      abline(sort(log(dfv2.out$Kd[dfv2.out$s==0]))[9900*0.999],0)
      text(5,max(log(dfv2.out$Kd), na.rm=TRUE)*0.95, "C", cex=2)
      text(30,max(log(dfv2.out$Kd), na.rm=TRUE)*0.95,round(getEmpPower(dfv2.out$Kd,dfv2.out$s_high==0),2))

    plot(ind, log(dfv2.out$Nd), col=col, pch=pch, ylab= "log(Nearest neighbor)", cex=cex)
      abline(sort(log(dfv2.out$Nd[dfv2.out$s==0]))[9900*0.999],0)
      text(5,max(log(dfv2.out$Nd), na.rm=TRUE)*0.95, "D", cex=2)
      text(30,max(log(dfv2.out$Nd), na.rm=TRUE)*0.95,round(getEmpPower(dfv2.out$Nd,dfv2.out$s_high==0),2))
}#end funciton

makeplot(dfv3.out) # raw

makeplot(dfv5.out) # MCD cov

makeplot(dfv6.subset) # MCD cov and "core" points

```

```{r}
# compare results
par(mfrow=c(1,1), mar=c(4,4,4,1))

plot(log(dfv3.out$Md), log(dfv5.out$Md), pch=pch, col=col, main="Md using all data vs. using MCD cov and mean")
abline(0,1, col="black", lwd=2)

plot(log(dfv2.out$Hd), log(dfv5.out$Hd), pch=pch, col=col,
     main = "Hd using all data vs MCD estimate")
abline(0,1, col="black", lwd=2)
plot(log(dfv5.out$Hd), log(dfv6.subset$Hd), pch=pch, col=col,
     main = "Hd using MCD cov vs using\nMCD cov and robust points")
abline(0,1, col="black", lwd=2)

plot(log(dfv5.out$Kd), log(dfv6.subset$Kd), pch=pch, col=col,
     main = "Kd using MCD cov vs using\nMCD cov and robust points")
abline(0,1, col="black", lwd=2)

plot(log(dfv5.out$Nd), log(dfv6.subset$Nd), pch=pch, col=col,
     main = "Nd using MCD cov vs using\nMCD cov and robust points")
abline(0,1, col="black", lwd=2)
```

