---
title: "Test Landsharc sims - make plots"
author: "Katie Lotterhos"
date: "May 24, 2016"
output: html_document
---

```{r setup, include=FALSE}
library(vioplot)
```
### Make summary plots


```{r}
summarydf <- read.table("results/LandsharcSummary3.txt", header=TRUE)
summarydf$demog <- factor(summarydf$demog, levels=c("IM", "IBD", "1R", "2R"))
#head(summarydf)
dim(summarydf)
# should be 72 rows
```

Compare power with and without MCD
```{r}
M <- summarydf$Md.mcd75.raw - summarydf$Md.default.raw
H <- summarydf$Hd.mcd75.raw - summarydf$Hd.default.raw
N <- summarydf$Nd.mcd75.raw - summarydf$Nd.default.raw
Hclust.ward <- summarydf$Hclust.wardD2.mcd75.raw-summarydf$Hclust.wardD2.default.raw
Hclust.ave <- summarydf$Hclust.Ave.mcd75.raw - summarydf$Hclust.Ave.default.raw
DCMS <- summarydf$DCMS.mcd75 - summarydf$DCMS.default

pdf("results/LandsharcComparePowerMCDminusPowerDefault.pdf", width=4, height=4)
  par(mar=c(4,4,1,1))
  boxplot(M, H, N, DCMS, names = c("Md", "Hd", "Nd", "DCMS"), las=2, col="grey90", ylab="(Power MCD) - (Power default)")
  abline(0,0, col="blue", lwd=3)
dev.off()

pdf("results/LandsharcComparePowerMCDminusPowerDefaultByDemog.pdf", width=4, height=8)
  par(mar=c(2,4,1,1), mfrow=c(4,1), oma=c(0,2,0,0))
  ylim= c(-0.4, 1)
  boxplot(M~summarydf$demog,  las=1, col="grey90", ylab="", ylim=ylim)
  abline(0,0, col="blue", lwd=3)
  text(0.5, 0.9, "Mahalanobis", adj=0)
  
  boxplot(H~summarydf$demog,  las=1, col="grey90", ylab="", ylim=ylim)
  abline(0,0, col="blue", lwd=3)
    text(0.5, 0.9, "Harmonic mean", adj=0)
  
    boxplot(N~summarydf$demog,  las=1, col="grey90", ylab="", ylim=ylim)
  abline(0,0, col="blue", lwd=3)
  text(0.5, 0.9, "Nearest Neighbor", adj=0)
  
      boxplot(DCMS~summarydf$demog,  las=1, col="grey90", ylab="", ylim=ylim)
  abline(0,0, col="blue", lwd=3)
  text(0.5, 0.9, "DCMS", adj=0)
  
  mtext("(Power MCD) - (Power default)", side=2, outer=TRUE, line=0)
dev.off()
```

```{r}
names(summarydf)
 stats <- c(7:10, 14:17,22, 23, 24, 25)
 names(summarydf)[stats]
 names2 <- c("BF", "rho", "XTX", "Z-score", "Md", "Hd", "Nd", "Hclust Ward",  "DCMS", "CSS", "Md-ranks", "Md-relative")

    colors <- c(rep("grey60", 4), rep("magenta",(length(stats)-4)))
      
  # addtext <- function(){
  #   text(2.5, 0.9, "Univariate", cex=1.1)
  #   text(6.5, 0.9, "Multivariate\nDefault", cex=1.1, col="lightblue")
  #   text(10.5, 0.9, "Multivariate\nMCD", cex=1.1, col="magenta")     
  #   text(14.5, 0.9, "Other", cex=1.1, col="darkblue") 
  # }
  
pdf("results/LandsharcSummary2.pdf", width = 15, height = 12)
    par(mfrow=c(4,1), mar=c(1,3,1,1),oma=c(1,3,1,0), cex=1.0)
    x <- 0.5
    y <- 1.1
   
    boxplot(summarydf[summarydf$demog=="IM",stats], col=colors,
            ylim=c(0,1.2), names=names2)
    text(x,y, "A. IM", cex=2, adj=0)
    #addtext()
    
    boxplot(summarydf[summarydf$demog=="IBD",stats], col=colors, ylim=c(0,1.2), names=names2)
    text(x,y, "B. IBD", cex=2)
    #addtext()
    
    boxplot(summarydf[summarydf$demog=="1R"  ,stats], col=colors,
            ylim=c(0,1.2), names=names2)
    text(x,y, "C. 1R", cex=2)
    #addtext()
    
    boxplot(summarydf[summarydf$demog=="2R"  ,stats], col=colors,
            ylim=c(0,1.2), names=names2)
    text(x,y, "D. 2R", cex=2)
    mtext("Empirical Power", side=2, outer=TRUE)
    #addtext()
    
dev.off()
```

Make plot showing robust points in 2 dimensions
```{r}
d1 <- read.table("data/KatieSims/2R_R30_1351142970_988_6_NumPops=30_NumInd=20Bayenv2LFMMpca.Cpval", header=TRUE)
  str(d1)
  d1$cols <- as.character(d1$cols)
  d1 <- d1[d1$SNPIncluded,]
  dfv <- d1[,c(1,3,4,5,10,12,13,15:17,34)]
  dim(dfv)
  str(dfv)
  names(dfv)
  X <- dfv[,c(9:10)]
  
  ### Change the colors
  d1$cols[d1$s_high==0.005] <- "lightblue"
  d1$cols[d1$s_high==0.01] <- "blue"
  d1$cols[d1$s_high==0.1] <- "darkblue"
  
plot1 <- function(){
  par(mfrow=c(1,1), mar=c(4,4,1,1))
  plot(X, col=rgb(0,0,0,0.05), pch=d1$pch1, xlim=c(-0.3,0.8), bty="l", ylim=c(20,80),
       xlab=expression("Spearman's "*rho),
       ylab=expression("X"^"T"*"X"))
  points(X[9901:9999,], col=d1$cols[9901:9999], pch=d1$pch1[9901:9999], ylim=c(20,80))
}
plot1()

library(cluster)
library(MASS)
library(rrcovNA)

### Add classical covariance
d2.99 <- qchisq(0.95, df = 2)

    C.ls <- cov(X) ; m.ls <- colMeans(X)
plot2 <- function(){
  lines(ellipsoidPoints(C.ls, d2.99, loc=m.ls), col="blue", lwd=2)
  text(0, 22, "All data", col="blue")
}
  plot1(); plot2()
  
### Add classical covariance of neutral loci
    C.ls2 <- cov(X[dfv$s_high==0,]) ; m.ls2 <- colMeans(X[dfv$s_high==0,])
plot3 <- function(){
  lines(ellipsoidPoints(C.ls2, d2.99, loc=m.ls2), col="black", lwd=2)
  text(0.22, 32, "Neutral only", col="black", adj=0)
}
  plot1(); plot2(); plot3()


### Add robust covariance to plot
  Cxy <- CovNAMcd(X, alpha=0.75) 
plot4 <- function(){
#cov.rob(X)
  #str(Cxy)
  #lines(ellipsoidPoints(Cxy$cov, d2 = d2.99, loc=Cxy$center), col="magenta", lwd=3)
  lines(ellipsoidPoints(Cxy@cov, d2 = d2.99, loc=Cxy@center), col="magenta", lwd=3)
  text(0.22, 29, "MCD estimate", col="magenta", adj=0)  
}
  plot1(); plot2(); plot3(); plot4()
   
### Add robust points to plot
pdf("results/VizCovariance.pdf", width = 6, height=5)  
  plot1(); plot2(); plot3(); plot4()
  points(X[Cxy@raw.wt==1,], pch=19, cex=0.1, col=rgb(1,0,1,0.2))  
  
  #points(mean(X[,1]), mean(X[,2]), pch=8, col="blue")
  #points(mean(X[d1$s_high==0,1]), mean(X[d1$s_high==0,2]), pch=8)
  
  levels(factor(d1$cols))

  legend(-0.25, 80, c("Neutral", "s = 0.005",
                     "s = 0.01","s = 0.1",
                     "Robust"), pch=c(19, 17, 17, 17, 20),
         col=c("#0000001A","lightblue", "blue",  "darkblue", "#FF00FF"), bty="n")
dev.off()
```

Make plot showing average signals for same dataset
```{r}
d2 <- read.table("results/landsharc/2R_R30_1351142970_988_6_NumPops=30_NumInd=20Bayenv2LFMMpca.Cpval.txt", header=TRUE)
str(d2)
d2$lev <- d2$s_high
d2$lev[d2$s_high==0 & !d2$IsRobust] = "Neutral\n(non-robust)"
d2$lev[d2$s_high==0 & d2$IsRobust] = "Neutral\n(robust)"
d2$lev <- factor(d2$lev, levels=c("Neutral\n(robust)", "Neutral\n(non-robust)", "0.005", "0.01", "0.1"))
col2 <- c("magenta", "black", "lightblue", "blue", "darkblue")

pdf("results/LandsharcAverageSignals.pdf", width=7, height=9)
  par(mfrow=c(2,1), mar=c(3,5,0,0))
  boxplot(d2$rho~d2$lev, col=col2, ylab=expression("Spearman's "*rho*" (Univariate)"))
  boxplot(log(d2$Md.mcd75.raw, 10)~d2$lev, col=col2,
        ylab="Log(Mahalanobis distance)\n(Multivariate)")
dev.off()

boxplot((summarydf$Md.mcd75.raw.relative-summarydf$Md.mcd75.raw)~summarydf$demog)
abline(0,0, col="blue")
```

Make plot comparing multivariate location and scatter - default vs. MCD vs. neutral
```{r}
mv <- read.csv("results/covsummaryLandsharc.csv", header=TRUE)
mv <- mv[-1,]
head(mv)

demog <- sapply(strsplit(as.character(mv$filename), "_"), function(x){x[1]})
demog <- factor(demog, levels=c("IM", "IBD", "1R","2R"))
(demoglevs <- levels(factor(demog)))

### Compare MCD and PCS
cond <- mv$Comparison1=="Neutral" & mv$Comparison2!="All"
mv3 <- mv[cond,]
levels(factor(mv3$Comparison2))
mv3$Comparison2 <- factor(mv3$Comparison2)

pdf("results/LandsharcComparetoNeutlocationscatterPCSvsMCD.pdf", width=5, height=7)
  par(mar=c(1,6,0.5,0.5), oma=c(4, 0,0,0), mfrow=c(2,1))
  boxplot(sqrt(mv3$ScatterDiff)~mv3$method*mv3$alpha, ylab="Mean difference with\nneutral multivariate scatter", las=1, col=c("magenta", "grey"), names=rep("",6))
  
  boxplot(sqrt(mv3$LocationDiff)~mv3$method*mv3$alpha, ylab="Mean difference with\nneutral multivariate location", las=1, col=c("magenta", "grey"),
          names=rep(c("MCD", "PCS"),3))
  mtext("alpha",1,at = 0.5, line=2.5)
  mtext("0.51",1,at = 1.5, line=2.5)
  mtext("0.75",1,at = 3.5, line=2.5)
  mtext("0.90",1,at = 5.5, line=2.5)
dev.off()

### Compare MCD only with Neutral
mv$method <- as.character(mv$method)
mv$method[is.na(mv$method)]="All"
cond <- mv$Comparison1=="Neutral" & mv$method!="PCS"
mv2 <- mv[cond,]
levels(factor(mv2$Comparison2))
mv2$Comparison2 <- factor(mv2$Comparison2)
mv2$Comparison2 <- factor(mv2$Comparison2, levels=c("MCD 0.51", "MCD 0.75", "MCD 0.90", "All"))

pdf("results/LandsharcComparetoNeutlocationscatter.pdf", height=7, width=5)
par(mar=c(1,5,0.5,0.5), mfrow=c(2,1), oma=c(3,0,0,0))

  boxplot(sqrt(mv2$ScatterDiff)~mv2$Comparison2, ylab="Mean difference with\nneutral multivariate scatter", las=2, names=c("","","",""), col=c(rep("magenta",3),"white"))

  boxplot(sqrt(mv2$LocationDiff)~mv2$Comparison2, ylab="Mean difference with\nneutral multivariate location", las=1, col=c(rep("magenta",3),"white"))
dev.off()

### Compare with All
cond <- mv$Comparison1=="All"
mv2 <- mv[cond,]
mv2$Comparison2 <- factor(mv2$Comparison2, levels=c("MCD 0.75", "MCD 0.99", "Neutral"))

pdf("results/LandsharcComparetoALLlocationscatter.pdf", height=8, width=5)
par(mar=c(3,5,1,0), mfrow=c(2,1))

  boxplot(sqrt(mv2$ScatterDiff)~mv2$Comparison2, ylab="Mean difference with\nmultivariate scatter of all data")

  boxplot(sqrt(mv2$LocationDiff)~mv2$Comparison2, ylab="Mean difference with\nmultivariate location of all data")
dev.off()  
```


Make plot of proportion of selected loci identified as robust
```{r}
head(summarydf)
attach(summarydf)

tapply(summarydf$ProportionS0.005MCDRobust, list(summarydf$demog, summarydf$alpha), mean)
tapply(summarydf$ProportionS0.01MCDRobust, list(summarydf$demog, summarydf$alpha), mean)
tapply(summarydf$ProportionS0.1MCDRobust, list(summarydf$demog, summarydf$alpha), mean)

# For IBD, the data set included four strengths of selection in each demography at the following percentages: sL = 0.001 (40% of the loci), sL = 0.005 (30%), sL = 0.01 (20%) and sL = 0.1 (10%). For IM, 1R and 2R, the data set included three strengths of selection at the following percentages: sL = 0.005 (50%), sL = 0.01 (33%) and sL = 0.1 (17%).

MakeExpectPlot <- function(demo){
    # demo is a character of the demography ("IM")
    boxplot(summarydf$ProportionS0.005MCDRobust[summarydf$demog==demo]~summarydf$alpha[summarydf$demog==demo], las=1, ylim=c(0,0.006))
    abline(100/10000*0.5,0, col="blue", lwd=2, lty=2)
    mtext(demo, side=2, line=4, las=1)
    if(demo=="IM"){
      mtext("s=0.005", side=3,  las=1)
    }
    boxplot(summarydf$ProportionS0.01MCDRobust[summarydf$demog==demo]~summarydf$alpha[summarydf$demog==demo], las=1, ylim=c(0,0.006))
    abline(100/10000*0.33,0, col="blue", lwd=2, lty=2)
    if(demo=="IM"){
      mtext("s=0.01", side=3, las=1)
    }
}

pdf("results/LandsharcProportionSelectedMCD.pdf", width=5, height=8)
  par(mfrow=c(4,2), oma=c(4,6,2,0), mar=c(2,4,0,0.1))
  MakeExpectPlot("IM")
      boxplot(summarydf$ProportionS0.005MCDRobust[summarydf$demog=="IBD"]~summarydf$alpha[summarydf$demog=="IBD"], las=1, ylim=c(0,0.004))
      abline(100/10000*0.3,0, col="blue", lwd=2, lty=2)
      mtext("IBD", side=2, line=4, las=1)
      boxplot(summarydf$ProportionS0.01MCDRobust[summarydf$demog=="IBD"]~summarydf$alpha[summarydf$demog=="IBD"], las=1, ylim=c(0,0.004))
      abline(100/10000*0.2,0, col="blue", lwd=2, lty=2)
  MakeExpectPlot("1R")
  MakeExpectPlot("2R")
  mtext("Proportion of selected loci identified as `robust` by MCD", side=2, outer=TRUE, line=4)
  mtext("Proportion of loci in dataset\nused for MCD", side=1, outer=TRUE, line=3)
dev.off()
```

Make plot of false positives if robust points are used as an empirical null distribution
```{r}
MakeFPPlot <- function(demo){
    # demo is a character of the demography ("IM")
    boxplot(summarydf$Md.mcd.subset.fp[summarydf$demog==demo]~summarydf$alpha[summarydf$demog==demo], las=1, ylim=c(0,0.25))
    mtext(demo, side=2, line=4, las=1)
}

  par(mfrow=c(4,1), oma=c(4,6,2,0), mar=c(2,4,0,0.1))
  MakeFPPlot("IM")
  MakeFPPlot("IBD")
  MakeFPPlot("1R")
  MakeFPPlot("2R")
```

All of the domgraphies have similar false positive rates. Make one plot for pub.

```{r}
pdf("results/LandsharcFalsePositivesMCD.pdf", width=3, height=3)
  par(mar=c(4,4,1,1), mfrow=c(1,1), oma=c(0,0,0,0))
  boxplot((summarydf$Md.mcd.subset.fp*9900)~summarydf$alpha, las=1, ylab="Number of false discoveries", xlab="Proportion of loci in dataset\nused for MCD")
  dev.off()
```

