---
title: "Dog data - calc stats"
author: "Katie Lotterhos"
date: "May 24, 2016"
output: html_document
---

### Load packages and data
```{r setup, include=FALSE}
#setwd("/Users/katie/Desktop/MINOTAUReval")
source("evalsims/getEmpiricalP3_BH.R")


if(!("devtools" %in% installed.packages())){install.packages("devtools", dependencies=TRUE)}
if(!("rrcovNA" %in% installed.packages())){install.packages("rrcovNA")}

library(qvalue)
library(rrcovNA)
library(devtools)
install_github("NESCent/MINOTAUR", ref="develop", force=TRUE)#, build_vignettes=TRUE)
library(MINOTAUR)
library(DMwR)

headdir <- "~/Desktop/MINOTAUReval/data/dog_full/"
outdir <- "~/Desktop/MINOTAUReval/data/dog_full_multvar_iHS_pi_TD_H12/"
#resultsdir <- "~/Desktop/MINOTAUReval/results/"
filelist <- list.files(headdir)
length(filelist)
if (!length(filelist)==25){print("Errr: missing simulation files")}

```

### Loop through each dataset
The following chunk of code does for each dataset:
1) calculate multivariate distances based on default covariance matrix
2) compares covariance matrix estimated from just neutral loci to covariance matrix on all data and MCD estimate of covariance
3) calculate multivariate distances based on MCD covariance matrix


```{r loop1}
cov_summary <- data.frame(filelist,Sall_Smcd0.75_cov=NA, Sall_Smcd0.75_mean=NA, Sall_Smcd0.90_cov=NA, Sall_Smcd0.90_mean=NA)
head(cov_summary)
for (i in 1:length(filelist)){
   print(c(i, filelist[i]))
  infile <- paste(headdir, filelist[i], sep="")
  dat <- read.csv(infile, header=TRUE)
  head(dat)
   dat2 <- dat[complete.cases(dat),]# #don't use complete cases if want to keep all data
  cols <- which(names(dat) %in% c("unstd_iHS", 
                                  "win51_pi",
                                  "win51_TajD",
                                  "win51_H12"))
  # Covariance of all loci
    Sall <- cov(dat2[complete.cases(dat2),cols])
    Sall
  # Covariance of MCD 75
    S_mcd75 <- CovNAMcd(dat2[,cols], alpha=0.75)
    #S_mcd75
  # Covariance of MCD 99
    S_mcd90 <- CovNAMcd(dat2[,cols], alpha=0.90)
    #S_mcd99    
    
  ## Mean difference between covmats
    Sall_mean <- colMeans(dat2[complete.cases(dat2),cols])
    
    cov_summary$Sall_Smcd0.75_cov[i] <- mean(abs(Sall-S_mcd75@cov))
    cov_summary$Sall_Smcd0.75_mean[i] <- mean(abs( Sall_mean-S_mcd75@center))
    cov_summary$Sall_Smcd0.90_cov[i] <- mean(abs(Sall-S_mcd90@cov)) 
    cov_summary$Sall_Smcd0.90_mean[i] <- mean(abs(Sall_mean-S_mcd90@center))
    print(cov_summary[i,])
}
   write.csv(cov_summary, file = paste(outdir,"covsummary.csv", sep=""))
``` 



```{r loop2}    
for (i in c(7,11:length(filelist))){
  print(c(i, filelist[i]))
  infile <- paste(headdir, filelist[i], sep="")
  dat <- read.csv(infile, header=TRUE)
  #head(dat)
  cols <- which(names(dat) %in% c("unstd_iHS", 
                                  "win51_pi",
                                  "win51_TajD",
                                  "win51_H12"))
    # don't include H: makes cov matrix singular
  cols
  dat2 <- dat[complete.cases(dat),]# #don't use complete cases if want to keep all data
  
    
  dat.out <- dat2[,c(1:4,cols)] 
  cols2 <- which(names(dat.out) %in% c("unstd_iHS", 
                                  "win51_pi",
                                  "win51_TajD",
                                  "win51_H12"))
    
 ### MCD alpha = 0.75
    S_mcd0.75 <- CovNAMcd(dat.out[,cols2], alpha=0.75)
    dat.out$IsRobust <- S_mcd0.75@raw.wt==1
    
### Get multivariate distance stats based on 
### raw statistics and all points
    try(dat.out$Md.default.raw <- Mahalanobis(dat.out, column.nums = cols2))
    
    try(dat.out$Hd.default.raw <- harmonicDist(dat.out, column.nums = cols2))
    
    #Kd.ML <- kernelDeviance(dat.out, cols2, bandwidth = bw)
    #(bw.best <- bw[which(Kd.ML==min(Kd.ML))[1]])
    #dat.out$Kd.default.raw <- kernelDist(dat.out, column.nums = cols2, bandwidth = bw.best)
    
    try(dat.out$Nd.default.raw <- neighborDist(dat.out, column.nums = cols2))
    #head(dat.out)
    
  ### Get multivariate distance stats based on 
### raw statistics and MCD 0.75
    try(dat.out$Md.mcd75.raw <- Mahalanobis(dat.out, column.nums = cols2, S=S_mcd0.75@cov, M=S_mcd0.75@center))
    
    try(dat.out$Hd.mcd75.raw <- harmonicDist(dat.out, column.nums = cols2, S = S_mcd0.75@cov, subset = dat.out$IsRobust))
    
    #Kd.ML <- kernelDeviance(dat.out, cols2, S = S_mcd0.75@cov, bandwidth = bw,subset = dat.out$IsRobust)
    #(bw.best <- bw[which(Kd.ML==min(Kd.ML))[1]])
    #dat.out$Kd.mcd75.raw <- kernelDist(dat.out, column.nums = cols2, bandwidth = bw.best, S = S_mcd0.75@cov, subset = dat.out$IsRobust)
    
    try(dat.out$Nd.mcd75.raw <- neighborDist(dat.out, column.nums = cols2,  S = S_mcd0.75@cov, subset = dat.out$IsRobust))
    #head(dat.out)

    print("dist done")

 
    ### Get p-values for raw stats
    #plot(dat.out$unstd_iHS) #right tailed
    #plot(dat.out$win51_pi) #left tailed
    #plot(dat.out$win51_TajD) #left tailed
    #plot(dat.out$win51_H12) #right tailed
    names(dat.out)[cols2]
    p.ranks1 <- stat_to_pvalue(dat.out, cols2[c(1,4)], two.tailed=FALSE, right.tailed=TRUE)
     p.ranks2 <- stat_to_pvalue(dat.out, cols2[c(2:3)], two.tailed=FALSE, right.tailed=FALSE)
     
    p.ranks <- cbind(p.ranks1[,1], p.ranks2, p.ranks1[,2])
    names(p.ranks) <- names(dat.out)[cols2]

    try(dat.out$DCMS.default <- DCMS(dat.out, cols2, dfp=p.ranks))
    #boxplot(dat.out$DCMS.default~dat.out$s_high)
    
    try(dat.out$DCMS.mcd75 <- DCMS(dat.out, cols2, 
                          subset = dat.out$IsRobust, dfp=p.ranks))
    #boxplot(dat.out$DCMS.mcd75~dat.out$s_high)
    
    try(dat.out$CSS <- CSS(dat.out, cols2, two.tailed=FALSE, right.tailed=TRUE)) 
    
      minusLogP.ranks <- -log(p.ranks,10)
    
    names(minusLogP.ranks) <- paste(names(p.ranks), ".minusLogP.rank", sep="")
    str(minusLogP.ranks)
    
    # Calculate Md on p.ranks relative to 0 using robust estimation of cov
    S_mcd0.75.p <- CovNAMcd(minusLogP.ranks, alpha=0.75)
    dat.out$Md.rankp.mcd75 <- Mahalanobis(minusLogP.ranks, S= S_mcd0.75.p@cov, M=rep(0, ncol(minusLogP.ranks)))
      
  
  write.csv(dat.out, file = paste(outdir,"set1_", filelist[i], sep=""))
}
geterrmessage()

```

 ### Get hclust distance based on raw statistics (and all points)
    ### default Hclust with wardD2
    Hcd <- outliers.ranking(dat.out[,cols2], method = "sizeDiff",
                 method.pars = NULL,clus = list(dist = "euclidean",alg = "hclust",meth = "ward.D2"),power = 1, verb = F)
    
    dat.out$Hclust.wardD2.default.raw <- -log(1-Hcd$prob.outliers)
    #boxplot(dat.out$Hclust.wardD2.default.raw~dat.out$s_high)  
    
    ### Hclust with wardD2 and robust points
    Hcd.mcd <- outliers.ranking(data= dat.out[dat.out$IsRobust,cols2], test.data= dat.out[,cols2], method = "sizeDiff", method.pars = NULL, clus = list(dist = "euclidean",alg = "hclust", meth = "ward.D2"), power = 1, verb = F)
    
    dat.out$Hclust.wardD2.mcd75.raw <- -log(1-Hcd.mcd$prob.outliers)
    


```{r}
sessionInfo()
```
