---
title: "Eval Landsharc sims"
author: "Katie Lotterhos"
date: "May 24, 2016"
output: html_document
---
setwd("/Users/katie/Desktop/Teaching/201609 Fall Ecological and Conservation Genetics/")
Class.list <- read.csv("EEMB3465.class_list2016.csv")
head(Class.list)
sample(Class.list$First.Name[Class.list$Level=="undergrad"], 4)

### Load packages and data
```{r setup, include=FALSE}
#setwd("/Users/katie/Desktop/MINOTAUReval")
source("evalsims/ComparePlot2.R")
#source("evalsims/distanceFunctionsOther.R")
source("evalsims/getEmpiricalP3_BH.R")
#source("evalsims/Getdf.R")


if(!("devtools" %in% installed.packages())){install.packages("devtools", dependencies=TRUE)}
if(!("rrcovNA" %in% installed.packages())){install.packages("rrcovNA")}

library(qvalue)
library(rrcovNA)
library(devtools)
library(DMwR)
install_github("NESCent/MINOTAUR", ref='develop')#, build_vignettes=TRUE)
library(MINOTAUR)
library(FastPCS)

headdir <- "data/KatieSims/"
resultsdir <- "~/Desktop/MINOTAUReval/results/"
filelist <- list.files(headdir)
filelist[55]
filelist <- filelist[-55]
length(filelist)
if (!length(filelist)==72){print("Error: missing simulation files")}

sessionInfo()
```


### Loop through each dataset
The following chunk of code does for each dataset:
1) calculate multivariate distances based on default covariance matrix
2) compares covariance matrix estimated from just neutral loci to covariance matrix on all data and MCD/PCS estimate of covariance

```{r}
cov_summary <- data.frame(filename=NA, 
                Comparison1=NA, Comparison2=NA, 
                ScatterDiff=NA, LocationDiff=NA,
                alpha=NA, method=NA)
head(cov_summary)

for (i in 1:length(filelist)){
  print(c(i, filelist[i]))
  
  infile <- paste(headdir, filelist[i], sep="")
  dat <- read.table(infile, header=TRUE)
  dat2 <- dat[dat$UseSNP,]
  cols <- which(names(dat) %in% c("rho", "log.bf", "xtx", "TW.Zscore"))

  # Covariance of all loci
    (Sall <- cov(dat2[,cols]))
    (Sall_mean <- colMeans(dat2[,cols]))
    
  # Covariance of neutral loci
    (Sneut <- cov(dat2[dat2$s_high==0,cols]))
    (Sneut_mean <- colMeans(dat2[dat2$s_high==0,cols]))
        
  # Covariance of MCD 51
    S_mcd51 <- CovNAMcd(dat2[,cols], alpha=0.51)
    S_mcd51@cov
    S_mcd51@center
  # Covariance of MCD 75
    S_mcd75 <- CovNAMcd(dat2[,cols], alpha=0.75)
    S_mcd75@cov
    S_mcd75@center
    #S_mcd75
  # Covariance of MCD 90
    S_mcd90 <- CovNAMcd(dat2[,cols], alpha=0.90)
    S_mcd90@cov
    S_mcd90@center
    #S_mcd99    
    
  # Covariance of PCS 51
    S_pcs51 <- FastPCS(dat2[,cols], alpha=0.51)
    S_pcs51$cov
    S_pcs51$center
  # Covariance of PCS 75
    S_pcs75 <- FastPCS(dat2[,cols], alpha=0.75)
    S_pcs75$cov
    S_pcs75$center
    #S_mcd75
  # Covariance of PCS 90
    S_pcs90 <- FastPCS(dat2[,cols], alpha=0.90)
    S_pcs90$cov
    S_pcs90$center
    
  ### Comparisons with neutral #########
    cov_summary<- rbind(cov_summary, data.frame(
      filename=filelist[i],
      Comparison1= "Neutral",
      Comparison2= "All",
      ScatterDiff = mean((Sall-Sneut)^2),
      LocationDiff = mean((Sall_mean-Sneut_mean)^2),
      alpha=NA, method=NA
    ))
    ### MCD
    cov_summary<-rbind(cov_summary, data.frame(
      filename=filelist[i],
      Comparison1= "Neutral",
      Comparison2= "MCD 0.51",
      ScatterDiff= mean((Sneut-S_mcd51@cov)^2),
      LocationDiff= mean((Sneut_mean-S_mcd51@center)^2),
      alpha=0.51, method="MCD"
    ))
    cov_summary<-rbind(cov_summary, data.frame(
      filename=filelist[i],
      Comparison1= "Neutral",
      Comparison2= "MCD 0.75",
      ScatterDiff= mean((Sneut-S_mcd75@cov)^2),
      LocationDiff= mean((Sneut_mean-S_mcd75@center)^2),
      alpha=0.75, method="MCD"
    ))
    cov_summary<-rbind(cov_summary, data.frame(
      filename=filelist[i],
      Comparison1= "Neutral",
      Comparison2= "MCD 0.90",
      ScatterDiff= mean((Sneut-S_mcd90@cov)^2),
      LocationDiff= mean((Sneut_mean-S_mcd90@center)^2),
      alpha=0.9, method="MCD"
    ))
    ### PCS
    cov_summary<-rbind(cov_summary, data.frame(
      filename=filelist[i],
      Comparison1= "Neutral",
      Comparison2= "PCS 0.51",
      ScatterDiff= mean((Sneut-S_pcs51$cov)^2),
      LocationDiff= mean((Sneut_mean-S_pcs51$center)^2),
      alpha=0.51, method="PCS"
    ))
    cov_summary<-rbind(cov_summary, data.frame(
      filename=filelist[i],
      Comparison1= "Neutral",
      Comparison2= "PCS 0.75",
      ScatterDiff= mean((Sneut-S_pcs75$cov)^2),
      LocationDiff= mean((Sneut_mean-S_pcs75$center)^2),
      alpha=0.75, method="PCS"
    ))
    cov_summary<-rbind(cov_summary, data.frame(
      filename=filelist[i],
      Comparison1= "Neutral",
      Comparison2= "PCS 0.90",
      ScatterDiff= mean((Sneut-S_pcs90$cov)^2),
      LocationDiff= mean((Sneut_mean-S_pcs90$center)^2),
      alpha=0.90, method="PCS"
    ))
    
    ### Comparisons with All
    cov_summary<- rbind(cov_summary, data.frame(
      filename=filelist[i],
      Comparison1= "All",
      Comparison2= "Neutral",
      ScatterDiff = mean((Sall-Sneut)^2),
      LocationDiff = mean((Sall_mean-Sneut_mean)^2),
      alpha=NA, method=NA
    )) #this is redundant, for plotting only
    cov_summary<-rbind(cov_summary, data.frame(
      filename=filelist[i],
      Comparison1= "All",
      Comparison2= "MCD 0.75",
      ScatterDiff= mean((Sall-S_mcd75@cov)^2),
      LocationDiff= mean((Sall_mean-S_mcd75@center)^2),
      alpha=0.75, method="MCD"
    ))
    cov_summary<-rbind(cov_summary, data.frame(
      filename=filelist[i],
      Comparison1= "All",
      Comparison2= "MCD 0.90",
      ScatterDiff= mean((Sall-S_mcd90@cov)^2),
      LocationDiff= mean((Sall_mean-S_mcd90@center)^2),
      alpha=0.90, method="MCD"
    ))
}    
   write.csv(cov_summary, file = paste("results/covsummaryLandsharc.csv", sep=""))
``` 


### Loop through each dataset
Evaluate empirical power

stat_to_pvalue # <- function(dfv, column.nums=1:ncol(dfv), subset=1:nrow(dfv), two.tailed=TRUE, right.tailed=TRUE)
DCMS # <- function(dfv, column.nums=1:ncol(dfv), subset=1:nrow(dfv), S=NULL, dfp, column.nums.p=1:ncol(dfp))
CSS # <- function(dfv, column.nums=1:ncol(dfv), subset=1:nrow(dfv), two.tailed=TRUE, right.tailed=TRUE){

DCMS now takes a dfp data frame along with dfv, which could be output from stat_to_pvalue but it could come from anywhere.

```{r}

  for (i in 1:length(filelist)){
    #i=21
    print(c(i, filelist[i]))
    infile <- paste(headdir, filelist[i], sep="")
    outfile <- paste(resultsdir, filelist[i], sep="")
    dat <- read.table(infile, header=TRUE)
    cols1 <- which(names(dat) %in% c("rho", "log.bf", "xtx", "TW.Zscore"))
    dat2 <- dat[dat$UseSNP,c(1,3,4,5,6,7,8,9,10,11,12,13,cols1)]
    dat.out <- dat2
    cols2 <- which(names(dat.out) %in% c("rho", "log.bf", "xtx", "TW.Zscore"))
    head(dat.out)    
    bw <- c(seq(0.01,0.1,by=0.03),seq(0.2,1,by=0.2), seq(2,5,by=1))
      #Bandwidths to try for Kd ML
    
    ### MCD alpha = 0.75
    S_mcd0.75 <- CovNAMcd(dat.out[,cols2], alpha=0.75)
    dat.out$IsRobust <- S_mcd0.75@raw.wt==1
    sum(dat.out$IsRobust)/nrow(dat.out)
    
        
### Get multivariate distance stats based on 
### raw statistics and all points
    dat.out$Md.default.raw <- Mahalanobis(dat.out, column.nums = cols2)
    
    dat.out$Hd.default.raw <- harmonicDist(dat.out, column.nums = cols2)
    
    #Kd.ML <- kernelDeviance(dat.out, cols2, bandwidth = bw)
    #(bw.best <- bw[which(Kd.ML==min(Kd.ML))[1]])
    #dat.out$Kd.default.raw <- kernelDist(dat.out, column.nums = cols2, bandwidth = bw.best)
    
    dat.out$Nd.default.raw <- neighborDist(dat.out, column.nums = cols2)
    #head(dat.out)

      # Calculate the proportion of neutral/selected loci identified as robust (non-outliers)
      ProportionNeutralMCDRobust <- sum(dat2$s_high[S_mcd0.75@raw.wt==1]==0)/sum(S_mcd0.75@raw.wt==1)
      ProportionS0.005MCDRobust <- sum(dat2$s_high[S_mcd0.75@raw.wt==1]==0.005)/sum(S_mcd0.75@raw.wt==1)
      ProportionS0.01MCDRobust <- sum(dat2$s_high[S_mcd0.75@raw.wt==1]==0.01)/sum(S_mcd0.75@raw.wt==1)
      ProportionS0.1MCDRobust <- sum(dat2$s_high[S_mcd0.75@raw.wt==1]==0.1)/sum(S_mcd0.75@raw.wt==1)
    
### Get multivariate distance stats based on 
### raw statistics and MCD 0.75
    dat.out$Md.mcd75.raw <- Mahalanobis(dat.out, column.nums = cols2, S=S_mcd0.75@cov, M=S_mcd0.75@center)
    
    dat.out$Hd.mcd75.raw <- harmonicDist(dat.out, column.nums = cols2, S = S_mcd0.75@cov, subset = dat.out$IsRobust)
    
    #Kd.ML <- kernelDeviance(dat.out, cols2, S = S_mcd0.75@cov, bandwidth = bw,subset = dat.out$IsRobust)
    #(bw.best <- bw[which(Kd.ML==min(Kd.ML))[1]])
    #dat.out$Kd.mcd75.raw <- kernelDist(dat.out, column.nums = cols2, bandwidth = bw.best, S = S_mcd0.75@cov, subset = dat.out$IsRobust)
    
    dat.out$Nd.mcd75.raw <- neighborDist(dat.out, column.nums = cols2,  S = S_mcd0.75@cov, subset = dat.out$IsRobust)
    #head(dat.out)

    print("dist done")
    
### Get hclust distance based on raw statistics (and all points)
    ### default Hclust with wardD2
    Hcd <- outliers.ranking(dat.out[,cols2], method = "sizeDiff",
                 method.pars = NULL,clus = list(dist = "euclidean",alg = "hclust",meth = "ward.D2"),power = 1, verb = F)
    
    dat.out$Hclust.wardD2.default.raw <- -log(1-Hcd$prob.outliers)
    #boxplot(dat.out$Hclust.wardD2.default.raw~dat.out$s_high)
    
    ### Hclust with wardD2 and robust points
    Hcd.mcd <- outliers.ranking(data= dat.out[dat.out$IsRobust,cols2], test.data= dat.out[,cols2], method = "sizeDiff", method.pars = NULL, clus = list(dist = "euclidean",alg = "hclust", meth = "ward.D2"), power = 1, verb = F)
    
    dat.out$Hclust.wardD2.mcd75.raw <- -log(1-Hcd.mcd$prob.outliers)
        
    ### default Hclust with Average
    Hcd <- outliers.ranking(dat.out[,cols2], method = "sizeDiff",
                 method.pars = NULL,clus = list(dist = "euclidean",alg = "hclust",meth = "average"),power = 1, verb = F)
    
    dat.out$Hclust.Ave.default.raw <- -log(1-Hcd$prob.outliers)
    #boxplot(dat.out$Hclust.Ave.default.raw~dat.out$s_high)
    
    ### Hclust with Average and robust points
    Hcd.mcd <- outliers.ranking(data= dat.out[dat.out$IsRobust,cols2], test.data= dat.out[,cols2], method = "sizeDiff", method.pars = NULL, clus = list(dist = "euclidean",alg = "hclust", meth = "average"), power = 1, verb = F)
    
    dat.out$Hclust.Ave.mcd75.raw <- -log(1-Hcd.mcd$prob.outliers)
     #boxplot(dat.out$Hclust.Ave.mcd75.raw~dat.out$s_high)
    
### Get p-values for raw stats
    #plot(dat.out$rho) #right tailed
    #plot(dat.out$log.bf) #right tailed
    #plot(dat.out$TW.Zscore) #right tailed
    #plot(dat.out$xtx) #right tailed
    
    p.ranks <- stat_to_pvalue(dat.out, cols2, two.tailed=FALSE, right.tailed=TRUE)
    names(p.ranks) <- names(dat.out)[cols2]

    dat.out$DCMS.default <- DCMS(dat.out, cols2, dfp=p.ranks) 
    #boxplot(dat.out$DCMS.default~dat.out$s_high)
    
    dat.out$DCMS.mcd75 <- DCMS(dat.out, cols2, 
                          subset = dat.out$IsRobust, dfp=p.ranks)
    #boxplot(dat.out$DCMS.mcd75~dat.out$s_high)
    
    dat.out$CSS <- CSS(dat.out, cols2, two.tailed=FALSE, right.tailed=TRUE) 
    #par(mfrow=c(1,1))
    #boxplot(dat.out$CSS~dat.out$s_high)

    minusLogP.ranks <- -log(p.ranks,10)
    
    names(minusLogP.ranks) <- paste(names(p.ranks), ".minusLogP.rank", sep="")
    str(minusLogP.ranks)
    
    # Calculate Md on p.ranks relative to 0 using robust estimation of cov
    S_mcd0.75.p <- CovNAMcd(minusLogP.ranks, alpha=0.75)
    dat.out$Md.rankp.mcd75 <- Mahalanobis(minusLogP.ranks, S= S_mcd0.75.p@cov, M=rep(0, ncol(minusLogP.ranks)))
    #boxplot(dat.out$Md.rankp.mcd75~dat.out$s_high)
    # plot(minusLogP.ranks, pch=19, cex=0.2, col=rgb(0,0,0,0.3))
    
     # Calculate Md on raw relative to least significant value using robust estimation of cov
    mins <- apply(dat.out[,cols2], 2, min)
    mins[c(2,4)] <- 0 # rho and Z-score would be relative to 0
    dat.out$Md.mcd75.raw.relative <- Mahalanobis(dat.out, cols2, S= S_mcd0.75@cov, M=mins)
    
    if(i%%6==0){
    pdf(paste("results/landsharc/", filelist[i],".pdf", sep=""), width=6, height=6)
    par(mfrow=c(1,1))
    
    # Md raw vs. Md rank p-values
    plot(dat.out$Md.mcd75.raw,dat.out$Md.rankp.mcd75, col=as.numeric(as.factor(dat.out$s_high)))
    
    # Md raw vs. Md raw relative
    plot(log(dat.out$Md.mcd75.raw,10), log(dat.out$Md.mcd75.raw.relative,10), col=as.numeric(as.factor(dat.out$s_high)))
    
    # Md raw relative vs. Md rank p-values
    plot(dat.out$Md.mcd75.raw.relative,dat.out$Md.rankp.mcd75, col=as.numeric(as.factor(dat.out$s_high)))
    
    # DCMS vs. to Md rank p-values
   plot(dat.out$DCMS.mcd75, dat.out$Md.rankp.mcd75, col=as.numeric(as.factor(dat.out$s_high)))
   
   par(mfrow=c(4,4), mar=c(1,1,1,1))
   plot(minusLogP.ranks, pch=19, cex=0.2, col=as.numeric(as.factor(dat.out$s_high)))
   dev.off()
    }
      
    head(dat.out)
    names(dat.out)
    
    ### Eval power for univariate and multivariate stats
      
    df.ep <- dat.out[,c(13:16,18:ncol(dat.out))] 
      
    Power <- apply(df.ep, 2, getEmpPower,  dat.out$s_high==0)
    
    ## Write all output to data frame  
    dat.out <- data.frame(dat.out, minusLogP.ranks)
    write.table(dat.out, paste("results/landsharc/", filelist[i],".txt", sep=""), col.names = TRUE, row.names=FALSE)
    
    
    out <- data.frame(infile = as.character(infile),
               demog = as.character(dat.out$demog[1]),
               ProportionNeutralMCDRobust,
               ProportionS0.005MCDRobust,
               ProportionS0.01MCDRobust,
               ProportionS0.1MCDRobust
      )
    
      allout <- cbind(out, data.frame(t(Power)))
    
        
    if (i==1){
      write.table(allout, file = paste(resultsdir,"LandsharcSummary3.txt", sep=""), append=FALSE, col.names=TRUE, row.names=FALSE)
    }else{
        write.table(allout, file = paste(resultsdir,"LandsharcSummary3.txt",sep=""), append=TRUE, col.names=FALSE, row.names=FALSE) 
                  }
} # end loop through files
```
